### 2021-08-22 doc/thesis/chapters/dsp.tex
implementation begin
move around newlines to avoid single lines

Figure~\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as `multiplex signal' in the literature.
Several of these parts are described in the following paragraphs, according to \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard} and \cite{ElectronicsNotesFmTutorial}.

@@

The bandwidth for this difference-signal spans 15~kHz on either side of the subcarrier.
It is used to generate a stereo audio signal, in combination with the mono signal.
This process is explained into more detail in Chapter~\ref{sec:demodulation-of-broadcast-fm}.

@@

The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as `Asynchronous Half-Wave Envelope Detector' \cite{DSPRelatedEnvelopeDetection}.
It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
In analog implementations, the thresholding unit is a diode.
The cutoff frequency of the lowpass needs to be adapted to the envelope signals' maximum frequency.
It needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.\\

Different methods may be implemented for the envelope detector.
For example, an `Asynchronous Full-Wave Envelope Detector' can be used.
Therefor, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{DSPRelatedEnvelopeDetection}.
The envelope is followed more exactly, because of the higher power that is available in the signal, since the entire signal is taken into account, and not only the positive half, as in the previous half-wave method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Demodulation of Broadcast FM}
\label{sec:demodulation-of-broadcast-fm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stereo Audio}

>>>>

Figure~\ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as \textit{multiplex signal} in the literature.
Several of these parts are described in the following paragraphs, according to \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard} and \cite{ElectronicsNotesFmTutorial}.

@@

The bandwidth for this difference-signal spans 15~kHz on either side of the subcarrier.
It is used to generate a stereo audio signal, in combination with the mono signal.
This process is explained into more detail in Chapter~\ref{sec:demodulation-of-a-broadcast-fm-channel}.

@@

The method implemented in Figure~\ref{fig:time_domain_envelope_detect} is one of the most simple ones and is usually referred to as \textit{Asynchronous Half-Wave Envelope Detector} \cite{DSPRelatedEnvelopeDetection}.
It consists of a thresholding unit to remove negative values and a conventional lowpass filter.
In analog implementations, the thresholding unit is a diode.
The cutoff frequency of the lowpass needs to be adapted to the envelope signals' maximum frequency.
It needs to be able to follow the signal, but also sufficiently smooth out the rectified signal.\\

Different methods may be implemented for the envelope detector.
For example, an \textit{Asynchronous Full-Wave Envelope Detector} can be used.
Therefor, the previous method only needs to swap the thresholding unit with an unit that calculates the absolut value.
In that architecture, a full-wave rectification is performed on the signal, which significantly improves the envelope detection accuracy \cite{DSPRelatedEnvelopeDetection}.
The envelope is followed more exactly, because of the higher power that is available in the signal, since the entire signal is taken into account, and not only the positive half, as in the previous half-wave method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Demodulation of a Broadcast FM Channel}
\label{sec:demodulation-of-a-broadcast-fm-channel}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stereo Audio}
\label{subsec:demod_stereo_audio}

### 2021-08-22 doc/thesis/chapters/implementation.tex
move around newlines to avoid single lines

  \item \textbf{De-Emphasis Filter}\\
      Pre-emphasis in the transmitter and its counterpart, the de-emphasis in the receiver, is a strategy to minimize the noise that is induced due to effects in the transmission, by attenuating and amplifying certain frequency parts of the signal.
      Both processes can be implemented, by using filters in the transmitter, as well as the receiver, respectively.
      The processes are implemented in the Matlab model, but are disabled in the generation of the input file that is used for the system.
      Thus, a de-emphasis filter is not implemented in the hardware, as it is not required for the show-case of this thesis.\\

@@

      \begin{itemize}
        \item \textbf{Between the branches \textit{Recover LR Diff} and \textit{Recover Mono}:}\\
          The \textit{Mono} branch only contains a single LP filter, while the \textit{LR Diff} branch contains two filters, namely a BP and an LP filter.
          The latter branch's signal is therefore delayed, with respect to the \textit{Mono} branch.\\
          The issue here is, that both signals are added and substracted at the end of the processing chain, to compute the left and right channel signals, respectively.
          Thus, they need to be aligned in time at that point.
          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP in the \textit{LR Diff} branch.

@@

      To overcome the group delay issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.\\
  \item \textbf{Carrier Recovery}\\
      The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{subsec:demod_stereo_audio}.
      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
\end{itemize}

@@

The next section in the block diagram represents multiple processes in parallel.
The LR Diff part is shifted to 38~kHz in the spectrum.
This is done by modulating the signal with the 38~kHz subcarrier.
A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
It band-limits the spectrum of the LR~Diff part between 23 and 53~kHz.
The BP filter introduces a constant group delay of the signal.
In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\

@@

Summing up, the Matlab model is not implemented in fixed point arithmetic, but is implemented in a way, that is close enough to the hardware implementation.
Therefore, it can serve as a reference for the hardware implementation.

@@

The Matlab implementation also includes a comprehensive range of analysis functionalities, especially in terms of visual plots.
These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.\\

An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the signal without any error, as the output signal does not exactly match the original input signal.
However, the result shows a clear separation of the left and right channels, and resembles the original input very closely, which is good enough for this case study.\\

>>>>

  \item \textbf{De-Emphasis Filter}\\
      Pre-emphasis in the transmitter and its counterpart, the de-emphasis in the receiver, is a strategy to minimize the noise that is induced due to effects in the transmission, by attenuating and amplifying certain frequency parts of the signal.
      Both processes can be implemented, by using filters in the transmitter, as well as the receiver, respectively.
      The processes are implemented in the Matlab model, but are disabled in the generation of the input file that is used for the system.
      Thus, a de-emphasis filter is not implemented in the hardware, as it is not required for the show-case of this thesis.

@@

      \begin{itemize}
        \item \textbf{Between the branches \textit{Recover LR Diff} and \textit{Recover Mono}:}\\
          The \textit{Mono} branch only contains a single LP filter, while the \textit{LR Diff} branch contains two filters, namely a BP and an LP filter.
          The latter branch's signal is therefore delayed, with respect to the \textit{Mono} branch.\\
          The issue here is, that both signals are added and substracted at the end of the processing chain, to compute the left and right channel signals, respectively.
          Thus, they need to be aligned in time at that point.
          Consequently, a delay block needs to be inserted in the \textit{Mono} branch, to compensate the group delay of the BP in the \textit{LR Diff} branch.\\

@@

      To overcome the group delay issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.

  \item \textbf{Carrier Recovery}\\
      The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{subsec:demod_stereo_audio}.
      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.
\end{itemize}

@@

The next section in the block diagram represents multiple processes in parallel.
The LR Diff part is shifted to 38~kHz in the spectrum.
This is done by modulating the signal with the 38~kHz subcarrier.
A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
It band-limits the spectrum of the LR~Diff part between 23 and 53~kHz.
The BP filter introduces a constant group delay of the signal.
In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.

@@

Summing up, the Matlab model is not implemented in fixed point arithmetic, but is implemented in a way, that is close enough to the hardware implementation.
Therefore, it can serve as a reference for the hardware implementation.\\

@@

The Matlab implementation also includes a comprehensive range of analysis functionalities, especially in terms of visual plots.
These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.
An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the signal without any error, as the output signal does not exactly match the original input signal.
However, the result shows a clear separation of the left and right channels, and resembles the original input very closely, which is good enough for this case study.\\
